### 2015-07-20 shrink/analysis.tex
added abstract, small reference fixes in rest of thesis.
a finished version: jul 20, 11 AM

%!TEX root = shrink.tex
\section{Energy vs. response time model}
\label{sec:analysis}

@@

%For a qualitative understanding of the factors that determine the goodness of the energy-response time tradeoff, we present an analytical model of server consolidation in \cdc s. We focus on a Zipf content workload the analyze the effect of the skewness of the Zipf workload and the server-load-dependent queuing delay on the energy-response time tradeoff.

\begin{figure}
\centering
\includegraphics[scale=0.36]{figures/model.pdf}
\vspace{-1.5in}
\caption{[Left] An illustrative server utilization vs. response time curve. [Right] Corresponding energy-response time curve.}
\label{fig:model}
\end{figure}

\subsection{Single server}
\label{sec:singleserver}
Consider a single \cdc\ server serving a workload of content requests from end-users that is unchanging, i.e., the arrival rate, popularity distribution, and size distribution of content across requests is fixed. Let $m$ denote the cache miss rate at the server, i.e., the fraction of requests for which the server contacts the origin server. Assume that the power drawn by a server $p(u)$ is a linear function of its utilization $u$, or the ratio of the incoming load and the server's capacity \cite{mathew12}. In this model, an idle server's power consumption is a fraction $I$ of its peak power $P$, and the power consumed increases linearly from $IP$ to $P$ as the utilization increases from 0 to 1, i.e., $p(u) = (I + (1-I)u)P$.

The end-to-end user-perceived response time is assumed to be the sum of three components as also illustrated by the three curves in Figure \ref{fig:model} (left): (1) Mean server delay $f(u)$, assumed to be a convex, increasing function of the utilization $u$ $(0\leq u\leq1)$; (2) Server-to-origin delay $B$, which is constant but incurred only upon a cache miss; (3) Client-to-server delay $A$, a constant. Thus, the total end-to-end user-perceived delay is $f(u) + mB + A$. As shown in the figure, realistic utilization vs. server delay profile are typically somewhere in between a straight line where the delay increases linearly with the utilization and an $L$-shaped curve where the delay is zero for all values of utilization less than 1 and is infinity at 1.

@@

The simplistic model above can serve as a first-order approximation of a datacenter viewed as a single logical server as follows. Suppose the datacenter consists of $N$ homogeneous servers, each identical to the one above. Suppose the total incoming load is uniformly distributed across all $N$ servers resulting in a fixed utilization $U$ at each server, which implies a power consumption of $(I+(1-I)U)P$ per server. Thus, the total power consumed is $(I+(1-I)U)PN$, and the mean response time is $f(U) + mB + A$. We have implicitly assumed here that the miss rate $m$ remains the same as above even though each server is getting a sampled transformation of the original request distribution.

What is the impact of consolidating the datacenter to $n<N$ servers on the total power consumed and user-perceived response time? To answer the first part, we observe that the utilization at each server increases by a factor $N/n$, so its power consumption would increase to $(I+(1-I)UN/n)P$. Thus, the consolidated datacenter's total power consumption is given by $(nI + (1-I)UN)P$, and the corresponding energy use relative to the peak (the x-axis in Figure  \ref{fig:model} (right)) is 
\begin{eqnarray}
\textit{benefit}  = \frac{ (nI + (1-I)UN)P}{(I+(1-I)U)PN}
\label{eq:benefit}
\end{eqnarray}

Computing the corresponding end-to-end response time with the consolidation as above is nontrivial as it requires us to account also for the increase in cache miss rates. Assuming that shutting down servers results in a proportional decrease by a factor $N/n$ in the total available storage (an assumption that is natural for clusters of commodity PCs--the more common option in practice--but not for CDCs relying on network storage), we need to compute the mean miss rate $m'$ that in general would be lower than $m$. In our numerical examples below, we derive $m'$ using a characteristic-time approximation model for an LRU cache \cite{che2002hierarchical}. 
The response time is given by $f(UN/n) + m'B + A$, and the corresponding response time 
inflation is given by 
\begin{eqnarray}
\textit{cost}  = \frac{ (f(UN/n) + mB + A)}{  (f(U) + m'B + A)}.
\label{eq:cost}
\end{eqnarray}

We make two observations based on the expression for the response time inflation above. First, the more skewed (closer to L-shaped as opposed to linear) the server's utilization vs. delay profile is, the less noticeable the impact on response time as $f(UN/n) \approx f(U)$ unless $UN/n \to 1$. Second, the more skewed (e.g., a high Zipf exponent) the popularity distribution is, the less noticeable the impact on response time as $m' \approx m$ assuming that the consolidated storage also suffices to cache the small fraction of popular objects contributing to the overwhelming portion of hits. We numerically exemplify this second insight next.
%
%\begin{figure}[t]
%\centering
%\begin{subfigure}[b]{0.23\textwidth}
%\includegraphics[scale=0.4]{graphs/vary-k.pdf}
%	\caption{}
%	\label{fig:vary-k}
%        \end{subfigure}
%\begin{subfigure}[b]{0.20\textwidth}
%\includegraphics[scale=0.4]{graphs/vary-z.pdf}
%	\caption{}
%	\label{fig:vary-z}
%        \end{subfigure}
%\caption{Energy-response time tradeoff for (a) varying models of server delay for a Zipf workload ($\alpha$ = 1.0) and (b) workloads with varying Zipf exponents.}
%\end{figure}

\textbf{Numerical example:} We evaluate the evaluate the energy-response time tradeoff for Zipfian content popularity distributions. We assume that the server's utilization vs. delay profile is given by $f(u) = Cu^K$, $K>=1$ is a model parameter and $C$ is a constant delay. Other model parameters are as follows: each server has a capacity = 1 request/sec, total load $L$ = 15 requests/sec, latency from clients to servers $A$ = 10 ms, latency from servers to origin servers $B$ = 100 ms, server delay coefficient $C$ = 400 ms, idle power fraction $I$ = 0.5, total number of unit-sized content $M$ = 100 million, storage per server $S$ = 1 million, number of servers $N$ = 100.

\begin{wrapfigure}[13]{r}{0.21\textwidth}
  \begin{center}
    \includegraphics[scale=0.4]{graphs/vary-z.pdf}
  \end{center}
  \caption{Energy-response time tradeoff for workloads with varying Zipf exponents.}
    \label{fig:vary-z}
\end{wrapfigure}

%\begin{figure}[t]
%\centering
%\begin{subfigure}[b]{0.23\textwidth}
%\includegraphics[scale=0.4]{graphs/vary-k.pdf}
%	\caption{}
%	\label{fig:vary-k}
%        \end{subfigure}
%\begin{subfigure}[b]{0.20\textwidth}
%\includegraphics[scale=0.4]{graphs/vary-z.pdf}
%	\caption{}
%	\label{fig:vary-z}
%        \end{subfigure}
%\caption{Energy-response time tradeoff for (a) varying models of server delay for a Zipf workload ($\alpha$ = 1.0) and (b) workloads with varying Zipf exponents.}
%\end{figure}

%In Figure \ref{fig:vary-k} and Figure \ref{fig:vary-z}, the energy use is shown relative to the peak energy use with $N$ = 100 active servers, and response time inflation is shown relative to the least response time observed in each graph.

%We assume that the server's utilization vs. delay profile $f(u) = Cu^K$, wherein $K>=1$ is a model parameter and $C$ is a constant delay. As $K$ increases from 1 to $\infty$, the delay profile becomes increasingly skewed.

@@

%Figure \ref{fig:vary-k} presents results for three models of server queuing delay with parameters $k$ = 1, 3 and 9; the Zipf exponent of the workload is $\alpha$ = 1.0 in all cases. In our model, the hit rates are independent of the server delay model. Therefore, the differences in response time among the curves is entirely due to a change in the server delay. We observe that the curve with $k$ = 9 achieves a significantly better energy-response time tradeoff than the curve k = 1. Further, the curve with $k$ = 9 stays nearly flat as energy use reduces by 60\% over peak. The reason is that server delays remain a small fraction of response times even at high utilization for $k$ = 9, and as a result, significant energy savings can be achieved with a small response time inflation.

Figure \ref{fig:vary-z} presents results for workloads with Zipf exponent $\alpha$ = 1.0, 0.8 and 0; $\alpha = 0$ results in a uniform distribution. The energy use is shown relative to the peak energy use with $N$ = 100 active servers, and response time inflation is shown relative to the least response time in the graph. The server  delay model parameter is $K$ = 9 for all workloads which implies that the difference in response times among them is due to a difference in  miss rates. We observe that workloads with higher Zipf exponents achieve a better energy-response time tradeoff. The reason is that a higher Zipf exponent means that most of the hits result from a small fraction of objects, so reducing the available storage only results in a small increase in miss rates. 
%A key implication is that real workloads which commonly exhibit a Zipf exponent close to 1.0 \cite{fayazbakhsh2013less} can expect to achieve a good energy vs. response time tradeoff.

\textbf{Summary:}
Our simple expository model suggests that the energy vs. response time tradeoff is more favorable as the skew in the server's utilization vs. delay profile increases and the skew in the content popularity distribution increases. Admittedly, this simplistic model has several critical limitations as it (1) does not consider workload dynamics, (2) assumes a perfect load balance among servers, (3) ignores the overhead of coordination between servers, and (4) implicitly assumes that the utilization vs. response time behavior and the cache size vs. miss rate behavior can be approximated by treating the \cdc\ as a single logical server of the same total capacity, ignoring the network fabric entirely.

Thus, a natural question is what do real, achievable energy vs. response time inflation tradeoffs look like in \cdc s? 
Further, does an increase server load or an increase in cache miss rates cause a greater impact on response time? 
To answer these questions, we present the design and implementation of Shrink and evaluate the tradeoff using a real workload from a large \cdc\  in the next two sections.

>>>>

%!TEX root = shrink.tex
\section{Energy vs. user-perceived latency model}
\label{sec:analysis}

@@

%For a qualitative understanding of the factors that determine the goodness of the energy-latency tradeoff, we present an analytical model of server consolidation in \cdc s. We focus on a Zipf content workload the analyze the effect of the skewness of the Zipf workload and the server-load-dependent queuing latency on the energy-latency tradeoff.

\begin{figure}
\centering
\includegraphics[scale=0.36]{figures/model.pdf}
\vspace{-1.5in}
\caption{[Left] An illustrative server utilization vs. latency curve. [Right] Corresponding energy-latency curve.}
\label{fig:model}
\end{figure}

\subsection{Single server}
\label{sec:singleserver}
Consider a single \cdc\ server serving a workload of content requests from end users that is unchanging, i.e., the arrival rate, popularity distribution, and size distribution of content across requests is fixed. Let $m$ denote the cache miss rate at the server, i.e., the fraction of requests for which the server contacts the origin server. Assume that the power drawn by a server $p(u)$ is a linear function of its utilization $u$, or the ratio of the incoming load and the server's capacity \cite{mathew12}. In this model, an idle server's power consumption is a fraction $I$ of its peak power $P$, and the power consumed increases linearly from $IP$ to $P$ as the utilization increases from 0 to 1, i.e., $p(u) = (I + (1-I)u)P$.

The end-to-end user-perceived latency is assumed to be the sum of three components as also illustrated by the three curves in Figure \ref{fig:model} (left): (1) Mean server latency $f(u)$, assumed to be a convex, increasing function of the utilization $u$ $(0\leq u\leq1)$; (2) Server-to-origin latency $B$, which is constant but incurred only upon a cache miss; (3) Client-to-server latency $A$, a constant. Thus, the total end-to-end user-perceived latency is $f(u) + mB + A$. As shown in the figure, realistic utilization vs. server latency profile are typically somewhere in between a straight line where the latency increases linearly with the utilization and an $L$-shaped curve where the latency is zero for all values of utilization less than 1 and is infinity at 1.

@@

The simplistic model above can serve as a first-order approximation of a datacenter viewed as a single logical server as follows. Suppose the datacenter consists of $N$ homogeneous servers, each identical to the one above. Suppose the total incoming load is uniformly distributed across all $N$ servers resulting in a fixed utilization $U$ at each server, which implies a power consumption of $(I+(1-I)U)P$ per server. Thus, the total power consumed is $(I+(1-I)U)PN$, and the mean latency is $f(U) + mB + A$. We have implicitly assumed here that the miss rate $m$ remains the same as above even though each server is getting a sampled transformation of the original request distribution.

What is the impact of consolidating the datacenter to $n<N$ servers on the total power consumed and user-perceived latency? To answer the first part, we observe that the utilization at each server increases by a factor $N/n$, so its power consumption would increase to $(I+(1-I)UN/n)P$. Thus, the consolidated datacenter's total power consumption is given by $(nI + (1-I)UN)P$, and the corresponding energy use relative to the peak (the x-axis in Figure  \ref{fig:model} (right)) is 
\begin{eqnarray}
\textit{benefit}  = \frac{ (nI + (1-I)UN)P}{(I+(1-I)U)PN}
\label{eq:benefit}
\end{eqnarray}

Computing the corresponding end-to-end latency with the consolidation as above is nontrivial as it requires us to account also for the increase in cache miss rates. Assuming that shutting down servers results in a proportional decrease by a factor $N/n$ in the total available storage (an assumption that is natural for clusters of commodity PCs--the more common option in practice--but not for CDCs relying on network storage), we need to compute the mean miss rate $m'$ that in general would be lower than $m$. In our numerical examples below, we derive $m'$ using a characteristic-time approximation model for an LRU cache \cite{che2002hierarchical}. 
The latency is given by $f(UN/n) + m'B + A$, and the corresponding latency 
inflation is given by 
\begin{eqnarray}
\textit{cost}  = \frac{ (f(UN/n) + mB + A)}{  (f(U) + m'B + A)}.
\label{eq:cost}
\end{eqnarray}

We make two observations based on the expression for the latency inflation above. First, the more skewed (closer to L-shaped as opposed to linear) the server's utilization vs. latency profile is, the less noticeable the impact on latency as $f(UN/n) \approx f(U)$ unless $UN/n \to 1$. Second, the more skewed (e.g., a high Zipf exponent) the popularity distribution is, the less noticeable the impact on latency as $m' \approx m$ assuming that the consolidated storage also suffices to cache the small fraction of popular objects contributing to the overwhelming portion of hits. We numerically exemplify this second insight next.
%
%\begin{figure}[t]
%\centering
%\begin{subfigure}[b]{0.23\textwidth}
%\includegraphics[scale=0.4]{graphs/vary-k.pdf}
%	\caption{}
%	\label{fig:vary-k}
%        \end{subfigure}
%\begin{subfigure}[b]{0.20\textwidth}
%\includegraphics[scale=0.4]{graphs/vary-z.pdf}
%	\caption{}
%	\label{fig:vary-z}
%        \end{subfigure}
%\caption{Energy-latency tradeoff for (a) varying models of server latency for a Zipf workload ($\alpha$ = 1.0) and (b) workloads with varying Zipf exponents.}
%\end{figure}

\textbf{Numerical example:} We evaluate the evaluate the energy-latency tradeoff for Zipfian content popularity distributions. We assume that the server's utilization vs. latency profile is given by $f(u) = Cu^K$, $K>=1$ is a model parameter and $C$ is a constant latency. Other model parameters are as follows: each server has a capacity = 1 request/sec, total load $L$ = 15 requests/sec, latency from clients to servers $A$ = 10 ms, latency from servers to origin servers $B$ = 100 ms, server latency coefficient $C$ = 400 ms, idle power fraction $I$ = 0.5, total number of unit-sized content $M$ = 100 million, storage per server $S$ = 1 million, number of servers $N$ = 100.

\begin{figure}[t]
\centering
    \includegraphics[scale=0.4]{graphs/vary-z.pdf}

  \caption{Energy-latency tradeoff for workloads with varying Zipf exponents.}
    \label{fig:vary-z}
\end{figure}

%\begin{figure}[t]
%\centering
%\begin{subfigure}[b]{0.23\textwidth}
%\includegraphics[scale=0.4]{graphs/vary-k.pdf}
%	\caption{}
%	\label{fig:vary-k}
%        \end{subfigure}
%\begin{subfigure}[b]{0.20\textwidth}
%\includegraphics[scale=0.4]{graphs/vary-z.pdf}
%	\caption{}
%	\label{fig:vary-z}
%        \end{subfigure}
%\caption{Energy-latency tradeoff for (a) varying models of server latency for a Zipf workload ($\alpha$ = 1.0) and (b) workloads with varying Zipf exponents.}
%\end{figure}

%In Figure \ref{fig:vary-k} and Figure \ref{fig:vary-z}, the energy use is shown relative to the peak energy use with $N$ = 100 active servers, and latency inflation is shown relative to the least latency observed in each graph.

%We assume that the server's utilization vs. latency profile $f(u) = Cu^K$, wherein $K>=1$ is a model parameter and $C$ is a constant latency. As $K$ increases from 1 to $\infty$, the latency profile becomes increasingly skewed.

@@

%Figure \ref{fig:vary-k} presents results for three models of server queuing latency with parameters $k$ = 1, 3 and 9; the Zipf exponent of the workload is $\alpha$ = 1.0 in all cases. In our model, the hit rates are independent of the server latency model. Therefore, the differences in latency among the curves is entirely due to a change in the server latency. We observe that the curve with $k$ = 9 achieves a significantly better energy-latency tradeoff than the curve k = 1. Further, the curve with $k$ = 9 stays nearly flat as energy use reduces by 60\% over peak. The reason is that server latencies remain a small fraction of latencies even at high utilization for $k$ = 9, and as a result, significant energy savings can be achieved with a small latency inflation.

Figure \ref{fig:vary-z} presents results for workloads with Zipf exponent $\alpha$ = 1.0, 0.8 and 0; $\alpha = 0$ results in a uniform distribution. The energy use is shown relative to the peak energy use with $N$ = 100 active servers, and latency inflation is shown relative to the least latency in the graph. The server  latency model parameter is $K$ = 9 for all workloads which implies that the difference in latencies among them is due to a difference in  miss rates. We observe that workloads with higher Zipf exponents achieve a better energy-latency tradeoff. The reason is that a higher Zipf exponent means that most of the hits result from a small fraction of objects, so reducing the available storage only results in a small increase in miss rates. 
%A key implication is that real workloads which commonly exhibit a Zipf exponent close to 1.0 \cite{fayazbakhsh2013less} can expect to achieve a good energy vs. latency tradeoff.

\textbf{Summary:}
Our simple expository model suggests that the energy vs. latency tradeoff is more favorable as the skew in the server's utilization vs. latency profile increases and the skew in the content popularity distribution increases. Admittedly, this simplistic model has several critical limitations as it (1) does not consider workload dynamics, (2) assumes a perfect load balance among servers, (3) ignores the overhead of coordination between servers, and (4) implicitly assumes that the utilization vs. latency behavior and the cache size vs. miss rate behavior can be approximated by treating the \cdc\ as a single logical server of the same total capacity, ignoring the network fabric entirely.

Thus, a natural question is what do real, achievable energy vs. latency inflation tradeoffs look like in \cdc s? 
Further, does an increase server load or an increase in cache miss rates cause a greater impact on latency? 
To answer these questions, we present the design and implementation of Shrink and evaluate the tradeoff using a real workload from a large \cdc\  in the next two sections.

### 2015-07-20 shrink/background.tex
a finished version: jul 20, 11 AM

The server to handle a particular end user's request is determined by a \emph{load balancer}. Servers in a \cdc\ have local storage used for caching content.  When a request reaches a server, it serves the content from its cache, if the content is available there, or else, it may fetch the content from a \emph{peer} cache in the same \cdc. If the requested content is unavailable on the \cdc's servers, the request results in a \emph{cache miss}. On a cache miss, content is fetched from remote locations, either from another \cdc\ or from \emph{origin servers}, at a cost of an increased response time, and served to the end user.

%\textbf{How does a \cdc\ reduce end-user response time?}

A  \cdc\ reduces end-user response time by avoiding load hotspots on servers and increasing cache hit rates, both with the help of its load balancer. Load balancer distributes request among servers to avoid load hotspots, which keeps server-load-dependent delays small. Further, it serves a content's requests from a small number of servers, which results in a small number of content replicas and increases overall cache hit rates.

@@

Despite a large body of research on automated server and switch consolidation schemes showing significant potential for reducing energy use, the impact of consolidation on user-perceived response times is not well understood. 
%these schemes are rarely used in practice or used with conservative or manually configured settings if at all \TBD{example and reference needed}. An important reason for this disparity is that operators are wary of the impact of consolidation on user-perceived response times, and the relationship between the two is not well understood. 
For example, a common problem approach in the literature \cite{mathew12, Jain} is to focus on dynamically estimating spare resources, implicitly assuming that shutting down the spare resources will have little or no impact on response times. However, in reality, it is rarely the case that one can, say, shut down 25 out of 400 servers with negligible response time impact. The precise impact depends on the workload dynamics and whether the application is constrained by compute, memory, disk, and/or network capacity. In the following section, we present a simple model that sheds light on these issues specifically in the context of CDCs.

%Our discussion shows that reducing response time and saving energy is important to \cdc\ operators. But, reducing energy via consolidation can inflate response time due to reasons such as to an increase in server load and cache misses. Therefore, a key question is whether \cdc s can expect to achieve a good tradeoff between the energy use and response time. We address this question using an analytical model in the next section.

%
%(1) Server consolidation creates a tradeoff between energy use and response times, and this tradeoff depends on server load dependent queuing delays as well as skewness in workload.
%(2) There exists an interaction between server consolidation and load balancing, which must be accounted for to accurately quantify energy-performance tradeoff.
%(3) A \cdc\ network contributes a non-trivial fraction to overall energy use, and a network-aware server consolidation can help reduce network energy.

@@

{\em 1)} Prior efforts on server consolidation, including efforts that focus on \cdc s \cite{mathew12, mathew2014energy},  do not account for the fact that different content are stored in the different servers, and further do not explicitly evaluate the impact of the consolidation on response times. However, it is important to devise a load balancer that will coordinate the content placement and server shutdown such that the cache miss rate (and hence response times) does not increase significantly as a result of the shutdown. Consider the example shown in Figure \ref{fig:server-content}. Suppose that all four servers are active during normal utilization periods, but the two servers on the right are turned off during low utilization periods. Squares with same letters represent replicas of the same content. In the top row, content C and D become unavailable within the cluster when servers are turned off, and  an user requesting that content will experience a longer response time as it will have to be refetched from the origin. Whereas in the bottom row one copy of all content is still available within the cluster, resulting in lower response times.

>>>>

The server to handle a particular end user's request is determined by a \emph{load balancer}. Servers in a \cdc\ have local storage used for caching content.  When a request reaches a server, it serves the content from its cache, if the content is available there, or else, it may fetch the content from a \emph{peer} cache in the same \cdc. If the requested content is unavailable on the \cdc's servers, the request results in a \emph{cache miss}. On a cache miss, content is fetched from remote locations, either from another \cdc\ or from \emph{origin servers}, at a cost of an increased latency, and served to the end user.

%\textbf{How does a \cdc\ reduce end-user latency?}

A  \cdc\ reduces user-perceived latency by avoiding load hotspots on servers and increasing cache hit rates, both with the help of its load balancer. Load balancer distributes request among servers to avoid load hotspots, which keeps server-load-dependent delays small. Further, it serves a content's requests from a small number of servers, which results in a small number of content replicas and increases overall cache hit rates.

@@

Despite a large body of research on automated server and switch consolidation schemes showing significant potential for reducing energy use, the impact of consolidation on user-perceived latencies is not well understood. 
%these schemes are rarely used in practice or used with conservative or manually configured settings if at all \TBD{example and reference needed}. An important reason for this disparity is that operators are wary of the impact of consolidation on user-perceived latencies, and the relationship between the two is not well understood. 
For example, a common problem approach in the literature \cite{mathew12, Jain} is to focus on dynamically estimating spare resources, implicitly assuming that shutting down the spare resources will have little or no impact on latencies. However, in reality, it is rarely the case that one can, say, shut down 25 out of 400 servers with negligible latency impact. The precise impact depends on the workload dynamics and whether the application is constrained by compute, memory, disk, and/or network capacity. In the following section, we present a simple model that sheds light on these issues specifically in the context of CDCs.

%Our discussion shows that reducing latency and saving energy is important to \cdc\ operators. But, reducing energy via consolidation can inflate latency due to reasons such as to an increase in server load and cache misses. Therefore, a key question is whether \cdc s can expect to achieve a good tradeoff between the energy use and latency. We address this question using an analytical model in the next section.

%
%(1) Server consolidation creates a tradeoff between energy use and latencies, and this tradeoff depends on server load dependent queuing delays as well as skewness in workload.
%(2) There exists an interaction between server consolidation and load balancing, which must be accounted for to accurately quantify energy-performance tradeoff.
%(3) A \cdc\ network contributes a non-trivial fraction to overall energy use, and a network-aware server consolidation can help reduce network energy.

@@

{\em 1)} Prior efforts on server consolidation, including efforts that focus on \cdc s \cite{mathew12, mathew2014energy},  do not account for the fact that different content are stored in the different servers, and further do not explicitly evaluate the impact of the consolidation on latencies. However, it is important to devise a load balancer that will coordinate the content placement and server shutdown such that the cache miss rate (and hence latencies) does not increase significantly as a result of the shutdown. Consider the example shown in Figure \ref{fig:server-content}. Suppose that all four servers are active during normal utilization periods, but the two servers on the right are turned off during low utilization periods. Squares with same letters represent replicas of the same content. In the top row, content C and D become unavailable within the cluster when servers are turned off, and  an user requesting that content will experience a longer latency as it will have to be refetched from the origin. Whereas in the bottom row one copy of all content is still available within the cluster, resulting in lower latencies.

### 2015-07-20 shrink/conclusions.tex
a finished version: jul 20, 11 AM

%!TEX root = shrink.tex
\section{Conclusions}
Content datacenters used for storing and serving content to end-users are common today. A major barrier to  widespread adoption of server and network consolidation is \cdc\ operators' concern on the impact on SLAs or user-perceived response times. Our work takes a step towards addressing this concern by presenting a model to quantify the energy savings vs. response time inflation trade-off in \cdc s.  A key insight, supported via experiments, is that despite server consolidation, cache hit rates remain close to an unconsolidated datacenter,  which helps mitigate the impact of consolidation on user-perceived response times. We have designed and implemented \shrink, a system that leverages this tradeoff to yield significant energy savings while affecting user-perceived response times in a controlled manner. \shrink's novel network-aware server consolidation algorithm reduces network energy use by up to 42\% compared to network-unaware server consolidation schemes. \shrink, in experiments based on content access traces from an Akamai datacenter, reduced energy use by 35\% compared to a baseline scheme that keeps entire datacenter always on while increasing mean response time by 8\% over it. Overall, our findings encourage deployment of consolidation techniques to reduce \cdc\ energy use.

>>>>

%!TEX root = shrink.tex
\section{Conclusions}
Content datacenters used for storing and serving content to end-users are common today. A major barrier to  widespread adoption of server and network consolidation is \cdc\ operators' concern on the impact on SLAs or user-perceived latencies. Our work takes a step towards addressing this concern by presenting a model to quantify the energy savings vs. latency inflation trade-off in \cdc s.  A key insight, supported via experiments, is that despite server consolidation, cache hit rates remain close to an unconsolidated datacenter,  which helps mitigate the impact of consolidation on user-perceived latencies. We have designed and implemented \shrink, a system that leverages this tradeoff to yield significant energy savings while affecting user-perceived latencies in a controlled manner. \shrink's novel network-aware server consolidation algorithm reduces network energy use by up to 42\% compared to network-unaware server consolidation schemes. \shrink, in experiments based on content access traces from an Akamai datacenter, reduced energy use by 35\% compared to a baseline scheme that keeps entire datacenter always on while increasing mean latency by 8\% over it. Overall, our findings encourage deployment of consolidation techniques to reduce \cdc\ energy use.
